SVM
===

Linear SVM classified
---------------------
Let's generate some dat ain two dimensions, and make them a little separated

```{r,warning=FALSE}
set.seed(10111)
x=matrix(rnorm(40),20,2) #40 random numbers in two columns
y=rep(c(-1,1),c(10,10)) #repeat each number 10 times
x[y==1,]=x[y==1,]+1 #move the mean
plot(x,col=y+3,pch=19)

library(e1071)
dat=data.frame(x,y=as.factor(y))
svmfit=svm(y~.,data=dat,kernel='linear',cost=10,scale=FALSE)
print(svmfit)
plot(svmfit,dat)
```

Make a prettier plot on our own: 
```{r}
#75 by 75 grid
make.grid=function(x,n=100){
  grange=apply(x,2,range)
  x1=seq(from=grange[1,1],to=grange[2,1],length=n)
  x2=seq(from=grange[1,2],to=grange[2,2],length=n)
  expand.grid(X1=x1,X2=x2)
}

xgrid=make.grid(x)
ygrid=predict(svmfit,xgrid)
plot(xgrid,col=c('red','blue')[as.numeric(ygrid)],pch=20,cex=0.2)
points(x,col=y+3,pch=19)
points(x[svmfit$index,],pch=5,cex=2)
```


Extract linear coefficients that describe the boundary:
```{r}
beta=drop(t(svmfit$coefs)%*%x[svmfit$index,])
beta0=svmfit$rho
plot(xgrid,col=c("red","blue")[as.numeric(ygrid)],pch=20,cex=.2)
points(x,col=y+3,pch=19)
points(x[svmfit$index,],pch=5,cex=2)
abline(beta0/beta[2],-beta[1]/beta[2])
abline((beta0-1)/beta[2],-beta[1]/beta[2],lty=2)
abline((beta0+1)/beta[2],-beta[1]/beta[2],lty=2)
```

Nonlinear SVM
-------------
Instead we will run SVM on some data hwere a non-linear boundary is called for. We will use the mixture data from ESL

```{r}
load(url("http://www.stanford.edu/~hastie/ElemStatLearn/datasets/ESL.mixture.rda"))
names(ESL.mixture)
rm(x,y)
attach(ESL.mixture)
plot(x,col=y+1)
dat=data.frame(y=factor(y),x)
fit=svm(factor(y)~.,data=dat,scale=FALSE,kernel='radial',cost=5)
```

Grid values already available
```{r}
xgrid=expand.grid(X1=px1,X2=px2)
ygrid=predict(fit,xgrid)
plot(xgrid,col=as.numeric(ygrid),pch=20,cex=0.2)
points(x,col=y+1,pch=19)

func=predict(fit,xgrid,decision.values=TRUE)
func=attributes(func)$decision
xgrid=expand.grid(X1=px1,X2=px2)
ygrid=predict(fit,xgrid)
plot(xgrid,col=as.numeric(ygrid),pch=20,cex=0.2)
points(x,col=y+1,pch=19)
contour(px1,px2,matrix(func,69,99),level=0,add=TRUE)
contour(px1,px2,matrix(prob,69,99),level=0.5,add=TRUE,col='blue')
